---
title: "Arsenic data"
output: 
  html_document:
author: "Austin Martinez"
date: "2024-06-13"
editor_options:
  markdown:
    wrap: 72
---

## Background

#please give a brief summary of the goal of this data report. Mention the data source (USGS paper), and a 3-5 sentence overview of what is contained. Then let's move into variable summaries. 

```{r setup, include=FALSE}
#setwd("~/Desktop")   #Set working directory for your local version of the cloned github repo
setwd("C:/Users/jhoover/Documents/GitHub/coPlateauWaterQuality")
library(kableExtra)
arsenicdata = read.csv("./01_data/Model_input_definitions 1.csv")
selected_predictors <- c("rt_clast_u", "rt_meta", "rt_plut_qtz", "PRMS8110Re", "LP2")
data <- subset(arsenicdata, Predictor_name %in% selected_predictors)
```
## Geologic Predictor Variables
#provide summaries of geologic predictor variables here, examples may include rocks, soils and other geologic factors

**rt_clast_u**: Generalized lithology classified as CLAST_U: sediments
and sedimentary rocks. Clastic sediments/rocks of unknown or highly
variable clast sizes.

**rt_meta**:Generalized lithology classified as META: Metamorphic rocks.

**rt_plut_qtz**: Generalized lithology classified as PLUT_QTZ: Igneous,
generally felsic, quartz-rich plutonic rocks such as granitoids,
granite, granodacite. Lithology of plutonics and volcanics are split out
based on quartz content using the IUGS classification presented in
Hyndman's 'Petrology of Igneous and Metamorphic Rocks' second edition,
1985, page 33, fields 1-5.


## Hydrology Predictor Variables
#provide summaries of hydrologic predictor variables here, examples may include depth to groundwater, lateral hydrologic position, stream flow, etc, acquifer characteristics

**PRMS8110Re**: average annual recharge to groundwater.

**LP2**: Lateral position for fourth order streams


## Climate Predictor Variables
#provide summaries of hydrologic predictor variables here, examples may include precipitation, temperature, evapotranspiration

## Land use/land cover Predictor Variables
#if applicable, provide summaries of Land use/land cover predictor variables here


# Predictor variable descriptions

```{r descriptions, include=TRUE}
NameDescription <- data.frame(data$Predictor_name, data$Description)
kable(NameDescription)
```

```{r all data, include=TRUE}
kable(data)
```

# Predictor variable descriptions2





```{r NameMax, include=TRUE}
NameMax <- data.frame(data$Predictor_name, data$Max)
kable(NameMax)
```

```{r NameMin, include=TRUE}
NameMin <- data.frame(data$Predictor_name, data$Min)
kable(NameMin)
```
