---
title: "Arsenic data"
author: "Austin Martinez"
date: "2024-06-13"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

## Background

### Overview Statement

This dataset that is provided by the  U.S. Geological Survey (USGS) was compiled and published by Melissa A. Lombard in 2021. This data spans from 1970 to 2013. This data was originally sourced to aid in the development of models that estimate the probability of private well arsenic concentrations exceeding various thresholds (1, 5, and 10 micrograms per liter) throughout the conterminous United States. 

The purpose of retrieving this dataset is to develop a model that uses this data to predict which areas of the Colorado Plateau are most at risk for underground water contamination by arsenic.



## Geologic Predictor Variables
#provide summaries of geologic predictor variables here, examples may include rocks, soils and other geologic factors

- **rt_carb**: Generalized lithology classified as CARB
- **rt_clast_c**: Generalized lithology classified as CLAST_C
- **rt_clast_f**: Generalized lithology classified as CLAST_F
- **rt_clast_u**: Generalized lithology classified as CLAST_U
- **rt_meta**: Generalized lithology classified as META
- **rt_plut_qtz**: Generalized lithology classified as PLUT_QTZ
- **D3**: Devonian; stratified sequence
- **De**: Devonian eugeosynclinal deposits
- **DSe**: New England; Devonian and Silurian eugeosynclinal deposits
- **Kg**: Plutonic and intrusive rocks; Cretaceous granitic rocks
- **O2**: Ordovician; stratified sequence
- **Oe**: Ordovician, eugeosynclinal
- **PP4**: Upper Pennsylvanian; stratified sequence
- **Pzg1**: Lower Paleozoic granitic rocks
- **Pzg2**: Middle Paleozoic granitic rocks
- **Q**: Quaternary; stratified sequence
- **Qv**: Quaternary volcanic rocks
- **S2**: Silurian; stratified sequence
- **Se**: Silurian eugeosynclinal deposits
- **Te2**: Tertiary; Eocene; stratified sequence
- **Tm**: Tertiary; stratified sequence
- **Tmc**: Tertiary; Miocene continental deposits
- **Toc**: Tertiary; Oligocene continental deposits
- **Tp**: Tertiary; stratified sequence
- **Tpc**: Tertiary; Pliocene continental deposits
- **Tpv**: Tertiary; Pliocene volcanic rocks
- **Tr**: Triassic; stratified sequence
- **Txc**: Tertiary; Paleocene continental deposits
- **uK2**: Upper Cretaceous Austin and Eagle Ford Groups
- **uK3**: Upper Cretaceous Taylor Group
- **uK3b**: Upper Cretaceous Taylor Group
- **uK4**: Upper Cretaceous Navarro Group
- **uPz**: Upper Paleozoic; stratified sequence
- **Wgn**: PreCambrian; metamorphic rocks
- **as_c**: Arsenic concentration in the soil C horizon
- **be_c**: Beryllium concentration in the soil C horizon
- **bi_c**: Bismuth concentration in the soil C horizon
- **inorgc_c**: Inorganic carbon concentration in the soil C horizon
- **mo_c**: Molybdenum concentration in the soil C horizon
- **ni_c**: Nickel concentration in the soil C horizon
- **orgc_c**: Phosphorus concentration in the soil C horizon
- **p_c**: Organic carbon concentration in the soil C horizon
- **sb_c**: Antimony concentration in the soil C horizon
- **se_c**: Selenium concentration in the soil C horizon



## Hydrology Predictor Variables
#provide summaries of hydrologic predictor variables here, examples may include depth to groundwater, lateral hydrologic position, stream flow, etc, acquifer characteristics

- **LP2**: Lateral position for fourth order streams
- **PRMS8110Re**: Average annual recharge to groundwater
- **LP2**: Lateral position for fourth order streams
- **LP4**: Lateral position for second order streams
- **LP6**: Lateral position for sixth order streams
- **BFI**: Mean annual base flow index
- **uc_999**: Water
- **Percent_Ti**: County estimates of U.S. tile drainage

## Climate Predictor Variables
#provide summaries of hydrologic predictor variables here, examples may include precipitation, temperature, evapotranspiration

- **DMppt8110**: Average annual precipitation from 1981 to 2010


## Ecological Region Variables (idk if you wanna call this land cover Predictor Variables)

- **na_10.1**: Cold deserts ecoregion
- **na_11.1**: Mediterranean California ecoregion
- **na_12.1**: Western Sierra Madre piedmont ecoregion
- **na_13.1**: Upper Gila Mountains ecoregion
- **na_6.2**: Western Cordillera ecoregion
- **na_7.1**: Marine West Coast forest ecoregion
- **na_8.1**: Mixed wood plains ecoregion
- **na_8.2**: Central USA plains ecoregion
- **na_9.2**: Temperate prairies ecoregion
- **na_9.3**: West-Central semi-arid prairies ecoregion
- **na_9.4**: South Central semi-arid prairies ecoregion
- **na_9.5**: Texas-Louisiana coastal plain ecoregion
- **na_9.6**: Tamaulipas-Texas semiarid plain ecoregion

## Soil and Sediment Variables

- **uc_12**: Alluvial sediments
- **uc_221**: Coastal zone sediments
- **uc_312**: Eolian sediments
- **uc_321**: Eolian sediments
- **uc_322**: Eolian sediments
- **uc_331**: Eolian sediments
- **uc_421**: Glacial till sediments
- **uc_422**: Glacial till sediments
- **uc_431**: Glacial till sediments
- **uc_451**: Glaciofluvial ice-contact sediments
- **uc_620**: Colluvial and alluvial sediments
- **uc_811**: Proglacial sediments
- **uc_812**: Proglacial sediments
- **uc_821**: Proglacial sediments
- **uc_822**: Proglacial sediments
- **uc_910**: Residual materials developed in igneous and metamorphic rocks
- **uc_920**: Residual materials developed in sedimentary rocks
- **uc_940**: Residual materials developed in carbonate rocks
- **uc_960**: Residual materials developed in bedrock
- **uc_970**: Residual materials developed in bedrock

## Output Variables 

- **As3Cat**: Code to assign observed arsenic concentration to categorical variable
- **bas1**: Code to assign observed arsenic concentration to binary variable
- **bas10**: Code to assign observed arsenic concentration to binary variable
- **bas5**: Code to assign observed arsenic concentration to binary variable
- **spl1**: Code to indicate if the location was used in the boosted regression tree As 1 model training or testing dataset
- **spl10**: Code to indicate if the location was used in the boosted regression tree As 10 model training or testing dataset
- **spl3cat**: Code to indicate if the location was used in the random forest model training or testing dataset
- **spl5**: Code to indicate if the location was used in the boosted regression tree As 5 model training or testing dataset

# U.S. Geological Survey (USGS) Dataset

```{r setup, include=TRUE}
#setwd("/Users/austinmartinez/Documents/GitHub/coPlateauWaterQuality/01_data")
#setwd("C:/Users/jhoover/Documents/GitHub/coPlateauWaterQuality/01_data")
#arsenicdata = read.csv("Model_input_definitions 1.csv")
#kable(arsenicdata)
```


# Categorical Variables stats summary

```{r arsenic_catagorical_data, include=TRUE}
# library(kableExtra)  #only need to load the library once
#setwd("/Users/austinmartinez/Documents/GitHub/coPlateauWaterQuality/01_data")
setwd("C:/Users/jhoover/Documents/GitHub/coPlateauWaterQuality/01_data")
library(kableExtra)
arsenic_catagorical_data <- read.csv("Arsenic_catagorical_variables.csv")
# Extracting specific columns
catagorical_data <- arsenic_catagorical_data[, c("Predictor_name", "Category_levels", "Categorical_observation_amounts")]
kable(catagorical_data)
```

# Continuous Variables stats summary

```{r arsenic_continous_data, include=TRUE}
#library(kableExtra)
#setwd("/Users/austinmartinez/Documents/GitHub/coPlateauWaterQuality/01_data")
setwd("C:/Users/jhoover/Documents/GitHub/coPlateauWaterQuality/01_data")
arsenic_continous_data <- read.csv("Arsenic_continuos_variables.csv")
# Extracting specific columns
extracted_data <- arsenic_continous_data[, c("Predictor_name", "Min", "Max", "X25._Quartile", "Mean", "X75._Quartile",
                                  "Standard_Deviation")]
kable(extracted_data)
```
