---
title: "Arsenic data"
output: 
  html_document:
author: "Austin Martinez"
date: "2024-06-13"
editor_options:
  markdown:
    wrap: 72
---

## Background

#please give a brief summary of the goal of this data report. Mention the data source (USGS paper), and a 3-5 sentence overview of what is contained. Then let's move into variable summaries. 

```{r setup, include=FALSE}
#setwd("~/Desktop")   #Set working directory for your local version of the cloned github repo
setwd("C:/Users/jhoover/Documents/GitHub/coPlateauWaterQuality")
library(kableExtra)
arsenicdata = read.csv("./01_data/Model_input_definitions 1.csv")
selected_predictors <- c("rt_clast_u", "rt_meta", "rt_plut_qtz", "PRMS8110Re", "LP2")
data <- subset(arsenicdata, Predictor_name %in% selected_predictors)
```
## Geologic Predictor Variables
#provide summaries of geologic predictor variables here, examples may include rocks, soils and other geologic factors

**rt_clast_u**: Generalized lithology classified as CLAST_U: sediments
and sedimentary rocks. Clastic sediments/rocks of unknown or highly
variable clast sizes.

**rt_meta**:Generalized lithology classified as META: Metamorphic rocks.

**rt_plut_qtz**: Generalized lithology classified as PLUT_QTZ: Igneous,
generally felsic, quartz-rich plutonic rocks such as granitoids,
granite, granodacite. Lithology of plutonics and volcanics are split out
based on quartz content using the IUGS classification presented in
Hyndman's 'Petrology of Igneous and Metamorphic Rocks' second edition,
1985, page 33, fields 1-5.

### Variables I added

**rt_carb**: Generalized lithology classified as CARB
**rt_clast_c**: Generalized lithology classified as CLAST_C
**rt_clast_f**: Generalized lithology classified as CLAST_F
**rt_clast_u**: Generalized lithology classified as CLAST_U
**rt_meta**: Generalized lithology classified as META
**rt_plut_qtz**: Generalized lithology classified as PLUT_QTZ
**D3**: Devonian; stratified sequence
**De**: Devonian eugeosynclinal deposits
**DSe**: New England; Devonian and Silurian eugeosynclinal deposits
**Kg**: Plutonic and intrusive rocks; Cretaceous granitic rocks
**O2**: Ordovician; stratified sequence
**Oe**: Ordovician, eugeosynclinal
**PP4**: Upper Pennsylvanian; stratified sequence
**Pzg1**: Lower Paleozoic granitic rocks
**Pzg2**: Middle Paleozoic granitic rocks
**Q**: Quaternary; stratified sequence
**Qv**: Quaternary volcanic rocks
**S2**: Silurian; stratified sequence
**Se**: Silurian eugeosynclinal deposits
**Te2**: Tertiary; Eocene; stratified sequence
**Tm**: Tertiary; stratified sequence
**Tmc**: Tertiary; Miocene continental deposits
**Toc**: Tertiary; Oligocene continental deposits
**Tp**: Tertiary; stratified sequence
**Tpc**: Tertiary; Pliocene continental deposits
**Tpv**: Tertiary; Pliocene volcanic rocks
**Tr**: Triassic; stratified sequence
**Txc**: Tertiary; Paleocene continental deposits
**uK2**: Upper Cretaceous Austin and Eagle Ford Groups
**uK3**: Upper Cretaceous Taylor Group
**uK3b**: Upper Cretaceous Taylor Group
**uK4**: Upper Cretaceous Navarro Group
**uPz**: Upper Paleozoic; stratified sequence
**Wgn**: PreCambrian; metamorphic rocks
**as_c**: Arsenic concentration in the soil C horizon
**be_c**: Beryllium concentration in the soil C horizon
**bi_c**: Bismuth concentration in the soil C horizon
**inorgc_c**: Inorganic carbon concentration in the soil C horizon
**mo_c**: Molybdenum concentration in the soil C horizon
**ni_c**: Nickel concentration in the soil C horizon
**orgc_c**: Phosphorus concentration in the soil C horizon
**p_c**: Organic carbon concentration in the soil C horizon
**sb_c**: Antimony concentration in the soil C horizon
**se_c**: Selenium concentration in the soil C horizon



## Hydrology Predictor Variables
#provide summaries of hydrologic predictor variables here, examples may include depth to groundwater, lateral hydrologic position, stream flow, etc, acquifer characteristics

**PRMS8110Re**: average annual recharge to groundwater.

### Variables I added
**LP2**: Lateral position for fourth order streams
**PRMS8110Re**: Average annual recharge to groundwater
**LP2**: Lateral position for fourth order streams
**LP4**: Lateral position for second order streams
**LP6**: Lateral position for sixth order streams
**BFI**: Mean annual base flow index
**uc_999**: Water
**Percent_Ti**: County estimates of U.S. tile drainage

## Climate Predictor Variables
#provide summaries of hydrologic predictor variables here, examples may include precipitation, temperature, evapotranspiration

### Variables I added
**DMppt8110**: Average annual precipitation from 1981 to 2010

## Land use/land cover Predictor Variables
#if applicable, provide summaries of Land use/land cover predictor variables here

## Ecological Variables (idk if you wanna call this land cover Predictor Variables)

### Variables I added
**na_10.1**: Cold deserts ecoregion
**na_11.1**: Mediterranean California ecoregion
**na_12.1**: Western Sierra Madre piedmont ecoregion
**na_13.1**: Upper Gila Mountains ecoregion
**na_6.2**: Western Cordillera ecoregion
**na_7.1**: Marine West Coast forest ecoregion
**na_8.1**: Mixed wood plains ecoregion
**na_8.2**: Central USA plains ecoregion
**na_9.2**: Temperate prairies ecoregion
**na_9.3**: West-Central semi-arid prairies ecoregion
**na_9.4**: South Central semi-arid prairies ecoregion
**na_9.5**: Texas-Louisiana coastal plain ecoregion
**na_9.6**: Tamaulipas-Texas semiarid plain ecoregion

## Soil and Sediment Variables

### Variables I added
**uc_12**: Alluvial sediments
**uc_221**: Coastal zone sediments
**uc_312**: Eolian sediments
**uc_321**: Eolian sediments
**uc_322**: Eolian sediments
**uc_331**: Eolian sediments
**uc_421**: Glacial till sediments
**uc_422**: Glacial till sediments
**uc_431**: Glacial till sediments
**uc_451**: Glaciofluvial ice-contact sediments
**uc_620**: Colluvial and alluvial sediments
**uc_811**: Proglacial sediments
**uc_812**: Proglacial sediments
**uc_821**: Proglacial sediments
**uc_822**: Proglacial sediments
**uc_910**: Residual materials developed in igneous and metamorphic rocks
**uc_920**: Residual materials developed in sedimentary rocks
**uc_940**: Residual materials developed in carbonate rocks
**uc_960**: Residual materials developed in bedrock
**uc_970**: Residual materials developed in bedrock

## Assigning/Indicating Code Variables (idk what else to name these)

### Variables I added
**As3Cat**: Code to assign observed arsenic concentration to categorical variable
**bas1**: Code to assign observed arsenic concentration to binary variable
**bas10**: Code to assign observed arsenic concentration to binary variable
**bas5**: Code to assign observed arsenic concentration to binary variable
**spl1**: Code to indicate if the location was used in the boosted regression tree As 1 model training or testing dataset
**spl10**: Code to indicate if the location was used in the boosted regression tree As 10 model training or testing dataset
**spl3cat**: Code to indicate if the location was used in the random forest model training or testing dataset
**spl5**: Code to indicate if the location was used in the boosted regression tree As 5 model training or testing dataset


# Predictor variable descriptions

```{r descriptions, include=TRUE}
NameDescription <- data.frame(data$Predictor_name, data$Description)
kable(NameDescription)
```

```{r all data, include=TRUE}
kable(data)
```

# Predictor variable descriptions2





```{r NameMax, include=TRUE}
NameMax <- data.frame(data$Predictor_name, data$Max)
kable(NameMax)
```

```{r NameMin, include=TRUE}
NameMin <- data.frame(data$Predictor_name, data$Min)
kable(NameMin)
```
